---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(gganimate)
library(kernlab)
```


# Visualisation

We need to come up with a function that takes a ternary matrix (SIR), a number of each state, and returns a colouring.
This should "look" spatially believable, where infected individuals are more likely to be clustered. 

Think of it as incrementing a matrix so that each additional infected or state change is a function of the previous state.
i.e. we have a discrete time step, and could conceivably generate a frame for each unique matrix state.

This might be achieved by a stochastic process / markov process that assigns probabilities for transitions, and randomly samples (with weights) from "candidate" cells to satisfy the change in SIR between each frame.

```{r}
init_matrix <- function(p, nrow, ncol, s0 = p, i0 = floor(p * 0.1)) {
  if (p != nrow * ncol) stop("population must be the product of nrow and ncol")
  
  mat <- matrix(NA_integer_, nrow = nrow, ncol = ncol) 
  # coords <- expand.grid(x = 1:ncol, y = 1:nrow)
  ids <- 1:(ncol*nrow)
  initial_infected <- sample(ids, i0)
  mat[initial_infected] <- 1L
  mat[-initial_infected] <- 0L
  mat
}
```

```{r}
plot_matrix <- function(mat) {
  p_data <- expand.grid(1:nrow(mat), 1:ncol(mat)) %>% 
    cbind(as.vector(mat)) %>% 
    set_names(c("y", "x", "state"))
  p <- ggplot(p_data, aes(x = x, y = y, fill = state)) +
    geom_raster() +
    coord_fixed()
  p
}
```

```{r, fig.width = 4}
mat <- init_matrix(640 * 640, 640, 640, i0 = 100)
plot_matrix(mat)
```

Now try tackling how to add more infected incrementally, based on location weighting (the hard bit).

```{r}
#' Given a state matrix, calculate the probability of each element
#' becoming infected in the next time step
#'
#' @param mat current state matrix
#' @param bandwidth width of kernel
#'
#' @return a matrix of
calc_infection_likelihood <- function(mat, bandwidth = 1, min_prob = 0.2) {
  H <- diag(c(bandwidth, bandwidth))
  
  infected <- which(mat == 1L, arr.ind = TRUE)
  kde <- ks::kde(
    x = infected, 
    H = H,
    xmin = 0L, xmax = 100L,
    eval.points = as.matrix(expand.grid(1:nrow(mat), 1:ncol(mat)))
  ) 
  
  # normalise output to range [min_prob, 1], assume fixed lower bound of 0
  # from kde output
  max_val <- max(kde$estimate)
  min_val <- min(kde$estimate)
  vec <- ((kde$estimate - min_val) / (max_val - min_val)) * 
    (1 - min_prob) + min_prob
  matrix(vec, nrow = nrow(mat))
}


increment_infected <- function(mat, n, kernel_fun) {
  infected <- which(mat == 1L, arr.ind = TRUE)
  
}
```

```{r}
infect_prob <- calc_infection_likelihood(mat, bandwidth = 10, min_prob = 0) 
plot_matrix(infect_prob)
```

```{r}
next_infected <- sample(1:length(mat), size = 1000, prob = infect_prob)
mat2 <- mat
mat2[next_infected] <- 1L
plot_matrix(mat2)
```

## SIR model component

```{r}
library(EpiDynamics)
```

```{r}
# Simulation parameters
infection_prob <- 0.08
recovery_rate <- 0.1

interactions_daily_normal <- 10
interactions_daily_distancing <- 3

n_days <- 150

# state intial conditions as a proportion of arbitrary total
I0 <- 1 / 1E4
initials <- c(S = 1 - I0, I = I0, R = 0)

run_sir_sim <- function(beta, gamma, initials, total_days) {
  sir <- SIR(
    pars = c(beta = beta, gamma = gamma), 
    init = initials,
    time = 1:total_days
  )
  sir
}
```




```{r}
# Parameters and initial conditions.
# parameters <- c(beta = 1.4247, gamma = 0.14286)
# initials <- c(S = 1 - 1e-06, I = 1e-06, R = 0)

# Solve and plot.
sir_no_distancing <- run_sir_sim(
  interactions_daily_normal * infection_prob, 
  recovery_rate,
  initials, 
  n_days
)

sir_distancing <- run_sir_sim(
  interactions_daily_distancing * infection_prob, 
  recovery_rate,
  initials, 
  n_days
)

no_distancing_results <- sir_no_distancing$results %>% 
  pivot_longer(
    cols = S:R,
    names_to = "variable",
    values_to = "value"
  ) %>% 
  mutate(scenario = "no_distancing")

distancing_results <- sir_distancing$results %>% 
  pivot_longer(
    cols = S:R,
    names_to = "variable",
    values_to = "value"
  ) %>% 
  mutate(scenario = "distancing")

p_data <- bind_rows(no_distancing_results, distancing_results) %>% 
  filter(variable == "I")

p <- ggplot(p_data, aes(x = time, y = value, colour = scenario)) +
  geom_line()

p
```

Calculate infected and recovered numbers, assuming fixed initial population size

```{r}
calc_state_frames <- function(sir_results, pop_size) {
  numbers <- sir_results %>% 
    mutate(
      S = round(S * pop_size),
      I = round(I * pop_size),
      R = pop_size - S - I
    )
  numbers
}
```


```{r}
calc_state_frames(sir_distancing$results, 640 * 640)
```









